---
title: "SARGEtools Example Markdown"
author: "Stephanie J. Wilson"
date: "Last Updated: 09-29-2026"
output:
  pdf_document:
    latex_engine: xelatex
---


This code is an example or template for creating a pdf summary of SARGE data. With this code you can plot data using the sarge_plot_defaults() function which will plot any data set as a default parameter 


# **Step 1: Set Up R to run this Code** 

## Download the SARGEtools package if you do not already have it
```{r download SARGEtools, message=FALSE, warning=FALSE}

#Use remotes to install the SARGEtools package from Github 

#if you don't have the remotes package use: 
    #install.packages("remotes")
library(remotes)

# Install from GitHub
remotes::install_github("Smithsonian/SARGEtools") 


```

## Read in packages that will be needed to run this code 
```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load required packages to run this markdown

library(tidyverse)
library(lubridate)
library(plotly)
library(DT)
library(SARGEtools)


```

## Set the time zone and date range of interest
```{r date time, message=FALSE, warning=FALSE}


# Set TIMEZONE that is relevant to your data 
TIMEZONE = "EST"


# Define the date window you want shown on the plots 
start_date = "2024-09-20"
end_date = "2024-09-27"


```


# **Step 2: Set your SARGE data configuration** 

## Define the data file descriptions 
### File name in example: file_descriptions.txt or file_descriptions.csv
### *User builds this file!* 

Need to set up a file that has links to your data files output by SARGE (these can be online in a Dropbox, Google Drive, Sharepoint, GitHub, etc). If you are not using an online repository, you can upload your files to a folder and put the path in the file_tab_descriptions.txt, but if you have an online repository you can put the link to the data. Note: If you are using google sheets, make the sheet so anyone with the link can view the file, then use the googlesheets package and add this line to your code: `googlesheets4::gs4_deauth()`

The file_descriptions will include user defined columns: 

 - **`data_type`**        = data type descriptor (e.g. Physiochemical, Hydrologic, GHG, Radon) need one for each data link, so typically one per instrument.
 
 - **`data_link`**        = location of the data (e.g. web link or file path).
 
 - **`skip`**             = how many lines will need to be skipped before reaching data in the file.
 
 - **`colname_override`** = these are the column names that we are defining for the linked file. 


*It is important to verify that each of these columns and rows match the data coming from the instrument of interest.*



## Data the column name descriptions file. This describes your variables, renames instrument or campbell code defined columns. 
### File name in example: colname_descriptions.csv
### *User builds this file!* 

The variable_descriptions file will include user defined columns: 

 - **`data_type`**        = data type descriptor (e.g. Physiochemical, Hydrologic, GHG, Radon) need one for each data link, should match file_descriptions data_types.
 
 - **`label`**            = label that will be used when plotting the data.
 
 - **`colname`**          = column name that will match one of the listed values in colname_override for the specified data_type.
 
 - **`color`**            = this is the R color that the variable will be plotted in.
 
 - **`default`**          = determines if this is a default plotting variable (T) or not (F); sarge_plot_defaults will only plot those listed as "T".
 
```{r configure data, warning=FALSE}

## Define your file_descriptions file and read it:  

# If you are using a text file: 
files_txt <- read_delim(
  #enter your path and file name: 
  "../template_markdown/configure_data/file_descriptions.txt",  
  
  show_col_types = FALSE, delim = "\t")

# If you are using a csv file: 
files_csv <- read_csv(
  #enter your path and file name: 
  "../template_markdown/configure_data/file_descriptions.csv",  
  
  show_col_types = FALSE)




# Define your colname_descriptions file and read it: 
variables <- read_csv(
  #enter your path and file name:
  "../template_markdown/configure_data/variable_descriptions.csv",   
  
  show_col_types = FALSE) %>%
  
  mutate(
    default = as.logical(default) ) %>%
  
  mutate(colname = if_else(is.na(colname), label, colname),
         
         label = if_else(is.na(label), colname, label))



```


# **Step 3: Read in your SARGE Data**

Use the sarge_read() function here to read in data listed in file_descriptions 

```{r read in data, warning=FALSE}

# Read in the data with sarge_read() function
data <- sarge_read(
  files = files_csv,       # dataframe containing your file_descriptions from Step 2
  variables = variables,   # dataframe containing your variable_descriptions from Step 2
  timezone = TIMEZONE,     # TIMEZONE value defined in Step 1
  resolution = "minute"    # time resolution you would like (e.g. "minute", "hour")
)

# Take a first look at the data 

head(data)     

```


# **Step 4: Plot the SARGE Data**

## Physiochemical Data: Measurements made by the Xylem EXO2 Sonde in the example data 

```{r plot physiochemical data, warning=FALSE}

# Set the desired data_type from your data (must match a data_type in variables)
data_type = "Physiochemistry"

# Plot the greenhouse gas data from SARGE 
exo_plot <- sarge_plot_defaults(
                    data = data,              #file_descriptions dataframe 
                    data_type = data_type,    #data_type you want to plot
                    start = start_date,       #the defined start date value
                    end = end_date,           #the defined end date value
                    variables = variables)    #variables dataframe with colnames, colors

# Print the Plot 
exo_plot 



# Plot the data for a shorter window of time and defined it in the function 
exo_plot1 <- sarge_plot_defaults(
                  data = data,                         #file_descriptions dataframe 
                  data_type = "Physiochemistry",       #data_type you want to plot
                  start = "2024-09-20 12:00:00",       #newly defined start date
                  end = "2024-09-20 14:00:00",         #newly defined end date 
                  variables = variables)               #variables dataframe with colnames, colors

# Print the Plot 
exo_plot1 

```    

## Hydrologic Data: Measurements made by the Sontek IQ ADCP in the example data 

```{r plot ADCP data, warning=FALSE}

# Set the desired data_type from your data (must match a data_type in variables)
data_type = "Hydrology"

# Plot the greenhouse gas data from SARGE 
adcp_plot <- sarge_plot_defaults(
                    data = data,              #file_descriptions dataframe 
                    data_type = data_type,    #data_type you want to plot
                    start = start_date,       #the defined start date value
                    end = end_date,           #the defined end date value
                    variables = variables)    #variables dataframe with colnames, colors

# Print the Plot 
adcp_plot 


```    

## Greenhouse Gas Data: Measurements made by a LICOR 7810 in the example data 

```{r plot GHG data, warning=FALSE}

# Set the desired data_type from your data (must match a data_type in variables)
data_type = "GHG"

# Plot the greenhouse gas data from SARGE 
ghg_plot <- sarge_plot_defaults(
                    data = data,              #file_descriptions dataframe 
                    data_type = data_type,    #data_type you want to plot
                    start = start_date,       #the defined start date value
                    end = end_date,           #the defined end date value
                    variables = variables)    #variables dataframe with colnames, colors

# Print the Plot 
ghg_plot 



# Plot the greenhouse gas data for a shorter window of time and defined it in the function 
ghg_plot1 <- sarge_plot_defaults(
                    data = data,                         #file_descriptions dataframe 
                    data_type = "GHG",                   #data_type you want to plot
                    start = "2024-09-20 12:00:00",       #newly defined start date
                    end = "2024-09-20 14:00:00",         #newly defined end date 
                    variables = variables)              #variables dataframe with colnames, colors

# Print the Plot 
ghg_plot1 

```    

## Radon Gas Data: Measurements made by a Durridge RAD7 in the example data 

```{r plot Radon data, warning=FALSE}


# Set the desired data_type from your data (must match a data_type in variables)
data_type = "Radon"

# Plot the greenhouse gas data from SARGE 
rad7_plot <- sarge_plot_defaults(
                    data = data,              #file_descriptions dataframe 
                    data_type = data_type,    #data_type you want to plot
                    start = start_date,       #the defined start date value 
                    end = end_date,           #the defined end date value
                    variables = variables)    #variables dataframe with colnames, colors

# Print the Plot 
rad7_plot 




```    








